---
title: "Which is the best country to live in?"
author: "Kirill Tumanov, Mattia Storero, Lucas Stark"
date: "14 12 2019"
output: 
  html_document:
    number_sections: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, include = F)
### Data 
library(OECD) #connection to OECD database
library(eurostat) #connection to eurostat database
library(wbstats) #connection 1 to world bank
library(WDI) #connection 2 or world bank
library(Quandl) #connection to Quandl
library(gapminder) #data from gapminder
### Cleaning and Visualizing
library(rvest) #if we need to harvest data from webpages
library(ggmap) #this way we can visualize data on maps (-> world map)
library(tidyverse) #basic cleaning/ organising functions
```

# Background
The idea of this project was created in middle of September 2019.
Three guys, from different lands had some discussions about theirs' countries.
Then, we started to think about particular question:
"Which is the world's best country to live in?"

To receive an answer for this attractive question we decided to select some 
variables, which were considerable for us.

The first significant indicator for us was health, which included: 
Life expectancy (gapminder)
Child mortality (gapminder)
Survival to age 65 by both genders (knoema)
Water Supply (knoema)

The second essential variable for us was Economics (Jobs/ Income), which included:
Inequality, gini coefficient (gapminder)
Income, GDP (gapminder)
Unemployment rate at the age of 25 - 54 (gapminder)

The third important indicator for us was Life Satisfaction, we used OECD database.

The fourth vital marker for us was Safety, which included:
Corruption Perception Index (gapminder)
Traffic deaths (gapminder)
Murders/100k (The world bank data)

The fifth considerable indicator for us was Education, which included:
Gender ratio of mean age in education (DATA BASE?)
Mean years in school (age group 25-54) (DATA BASE?)

The sixth substantial marker for us was Environment, we used CO2 emissions, tonnes per person (gapminder)

# Research Question
To create a well-developed project, we asked the following questions:
What is the current state of the indicator (2019 or the last year)?
How did it develop over the years?
How is it distributed over the continents?

# Data Sources
```{r import}
### 1 OECD ###
#- load table with all available variables
OECDdatasets <- get_datasets()
avg_unemployment <- get_dataset("AVD_DUR")

### 2 EUROSTAT ###
#- load table with all available variables
ES_datasets <- search_eurostat("")
#df <- get_eurostat()

### 3 WORLD BANK ###
#- load table with all available variables
WDI_datasets <- as_tibble(WDIcache()$series)
temp <- stringr::str_detect(WDI_datasets[["name"]], "net enrolment")
temp2 <- WDI_datasets[temp,]
#temp2[["description"]]
#load datasets:
WDI_ch_mor <- WDI(indicator = "SH.DYN.MORT.Q3")
WDI_ED_pri <- WDI(indicator = "UIS.NERT.1") # % in primary education
WDI_ED_sec <- WDI(indicator = "UIS.NERA.3") # % in upper secondary education

### 4 Quandl ###
#- a way to search for datasets:
Q_datasets <- Quandl.search("time in.* education")
# load datasets:
# df <- Quandl(code)
Q_lifeExp <- Quandl("22254767")
Q_x2 <- Quandl("")

### 5 Gapminder ###
#- 6 variables: country, continent, year, lifeExp, pop, gdpPercap
summary(gapminder)


```

# Data Cleaning
```{r cleaning}

```

# Data Analysis
```{r analysis}

```

## Health Variables

## Economic Variables

## Life Satisfaction variables

## Safety Variables

## Educational variables

## Environmental variables

# Conclussion


